SMT-Boosted Security Types for Low-Level MPC

### 2025-11-19 toplas26/metaty.tex
checkpoint

Share types rely on SMT entailments via the \TirName{ShareEntails} and
\TirName{PubEntails} rules, which are not syntax directed. Our goal
with the $\metaprot$ language is to automate share type checking. To
enable this we introduce \emph{share type annotations}. These
annotations allow programmer specification of properties that can be
verified as part of the Hoare triple analysis defined in Section
\ref{section-metahtrip} and that enable automated checking of types.

@@

\begin{lemma}
  Given $\prog_1,\Gamma_1,R_1 : {OK}$ and $\cmd \redx \prog_2$ and $\htrip{\eqs_1}{\cmd}{\eqs_2}$ with
  $\eqs_1$ satisfiable and $\toeq{\prog_1} \models \eqs_1 \wedge \toeq{\xdefs_1}$ and
  $\shtrip{\Gamma_1,R_1,\xdefs_1}{\cmd}{\Gamma_2,R_2,\xdefs_2}$.
  Then $\prog_1;\prog_2,\Gamma_2,R_2 : {OK}$ and $\toeq{\prog_1;\prog_2} \models \toeq{\xdefs_2}$.
\end{lemma}

\begin{theorem}
  Given $\htrip{\eqspre}{\ttt{main()}}{\eqs}$ and $\ttt{main} : \Gamma_1 \rightarrow (\Gamma_2,R)$
  and $\main() \redx \prog$ and $\progpre,\Gamma_1,R' : {OK}$ with $R' \cap R = \varnothing$
  and $\toeq{\progpre} \models \eqspre$. Then $(\progpre;\prog),\Gamma_2,R : {OK}$.
\end{theorem}

>>>>

Share types rely on SMT entailments via the \TirName{ShareEntails} and
\TirName{PubEntails} rules, which are not syntax directed, and likewise
the interactive typing judgement is nondeterministic. Our goal
with the $\metaprot$ language is to automate share type checking. To
enable this we introduce \emph{share type annotations}. These
annotations allow programmer specification of properties that can be
verified as part of the Hoare triple analysis defined in Section
\ref{section-metahtrip} and that enable automated checking of types.

@@

Demonstrating soundness of $\metaprot$ type checking is similar to $\minicat$--
we show the ${OK}$ invariant is preserved by the algorithm. By Theorem \ref{theorem-OK}
soundness follows. Lemma \ref{lemma-metaty} follows by appeal to Lemma \ref{lemma-sty-noninteractive-sound}
since the algorithmic judgement $\Vdash$ is sound with regard to the noninteractive
typing rules by definition. Theorem \ref{theorem-metaty} follows as a special
case of Lemma \ref{lemma-metaty}.
\begin{lemma}
  \label{lemma-metaty}
  Given $\prog_1,\Gamma_1,R_1 : {OK}$ and $\cmd \redx \prog_2$ and $\htrip{\eqs_1}{\cmd}{\eqs_2}$ with
  $\eqs_1$ satisfiable and $\toeq{\prog_1} \models \eqs_1 \wedge \toeq{\xdefs_1}$ and
  $\shtrip{\Gamma_1,R_1,\xdefs_1}{\cmd}{\Gamma_2,R_2,\xdefs_2}$.
  Then $\prog_1;\prog_2,\Gamma_2,R_2 : {OK}$ and $\toeq{\prog_1;\prog_2} \models \toeq{\xdefs_2}$.
\end{lemma}

\begin{theorem}
  \label{theorem-metaty}
  Given $\htrip{\eqspre}{\ttt{main()}}{\eqs}$ and $\ttt{main} : \Gamma_1 \rightarrow (\Gamma_2,R)$
  and $\main() \redx \prog$ and $\progpre,\Gamma_1,R' : {OK}$ with $R' \cap R = \varnothing$
  and $\toeq{\progpre} \models \eqspre$. Then $(\progpre;\prog),\Gamma_2,R : {OK}$.
\end{theorem}

Note that Theorem \ref{theorem-metaty} assumes the existence of
$\progpre$ with $\progpre,\Gamma_1,R' : {OK}$ and $R' \cap R =
\varnothing$ and $\toeq{\progpre} \models \eqspre$. The assumption of
some preprocessing protocol such as Beaver triple or MAC share
distribitions is typical, and our approach does not require the
protocol to be defined in detail, just (weakened) postconditions of it
that are sufficient to verify the protocol being programmed.

Theorem \ref{theorem-metaty} also does not provide a complete recipe
for obtaining noninterference modulo output-- as suggested by Theorem
\ref{theorem-sty}, it is also necessary to check closure of
$\Gamma_2$. There are various ways this can be accomplished for the
$\metaprot$ analysis-- for example, we evaluate $\ttt{main}$ to obtain
all program variables and then check closure of $\Gamma_2$ as
prescribed by Definition \ref{def-sty-closed}. We can also add checks
to the type system to guarantee that all messages exchanged between
clients are classified when type checking functions, which is
a simple addition.
